Local metal-insulator transition in a generalised impurity Anderson model

### 2022-05-04 gen-siam-manuscript/gen-siam.tex
added corectec figures

The Hubbard model is one of the fundamental models for strong electronic correlation; in its simplest form, it features a single band of conduction electrons hopping on a lattice and interacting via local correlations that provide a cost \(U\) if any site is doubly occupied. There are two trivial limits of the model. At \(U=0\), the bath consists of just a kinetic energy part, and the ground state is just a filled Fermi sea. At \(t=0\), each lattice site decouples from the rest and becomes a local moment, which under symmetry-breaking becomes a Neel antiferromagnet. This suggests that on increasing \(U/t\) beyond some critical value, the system might undergo a phase transition from a metallic state to an insulating state~\cite{Mott_1949}. This transition is reflected in the local spectral function - while it has a well-defined zero energy peak in the metallic phase, it is gapped in the insulating phase.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.5\textwidth]{../figures/dos_diff.pdf}
\caption{Two kinds of bath that an impurity can hybridise into. The left panel shows a non-interacting conduction band with a flat density of states. The right panel shows an interacting bath with an energy-dependent density of states. In the latter case, the impurity "feels" a reduced effective bandwidth defined by the width of the central peak.}
\end{figure}

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.45\textwidth]{../figures/zeromode_bare.pdf}
	\caption{While we have studied the full model under renormalisation group, often we will turn to a simplified zero-bandwidth version of the model that is obtained by ignoring the kinetic energy part of the Hamiltonian. This zero-bandwidth model is effectively a two site model.}
	\label{zeromode-bare}
\end{figure}

@@

\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.45\textwidth]{../figures/urg-scheme.pdf}
	\caption{High energy - low energy scheme defined and used in the URG method. The states away from the Fermi surface form the UV subspace and are decoupled first, leading to a Hamiltonian which is more block-diagonal and comprised of only the IR states near the Fermi surface.}
	\label{urg-scheme}
\end{figure}

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.48\textwidth]{../figures/J_Ub.pdf}
	\includegraphics[width=0.48\textwidth]{../figures/U_Ub.pdf}
	\includegraphics[width=0.48\textwidth]{../figures/V_Ub.pdf}
	\caption{Variation of couplings \(U,V\) and \(J\) along the RG transformations, for three values of the transition-tuning ratio \(r\). x-axis represents the RG steps, the rightmost point is the first RG step and leftmost point is the final RG step. The values of the bare couplings are given in the title. All values are in units of the half-bandwidth \(D_0\), which is chosen to be 10. The blue curves represent the flows for \(r > r_c\), where \(J\) is relevant. The purple curves represent the RG flows for \(r < r_c\), where \(J\) is irrelevant. The red curves are exactly at the critical point \(r = r_c\), where \(J\) is marginal.}
	\label{rg-flow}
\end{figure}

@@

\begin{figure}[!htb]
	\includegraphics[width=0.5\textwidth]{../figures/phase-map-MIT.pdf}
	\caption{Phase diagram of the GAIM, in the space of couplings \(U_0\) and \(J_0\). \(U_b\) is set to \(-U_0/10\) and \(V_0\) is set to \(2J_0\). The various phases \(A\) through \(D\) are characterised in the text.}
	\label{phase-diag}
\end{figure}

@@

\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.49\textwidth]{../figures/odlro.pdf}
	\caption{Growth in the pair-fluctuation terms \(c^\dagger_{k \uparrow}c^\dagger_{k \downarrow} c_{k^\prime \downarrow}c_{k^\prime \uparrow}\), as \(r \to 0.25\). This comes at the cost of the spin-flip fluctuations \(S_d^+ S_0^- + \text{h.c.}\), which decrease towards the transition. The inset plots the same quantities but for a larger range of \(r\), and the pair fluctuations are seen to pick up abruplty after the transition while the spin fluctuations vanish.}
	\label{pair_fluc}
\end{figure}

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.5\textwidth]{../figures/J_bandwidth.pdf}
	\caption{Increase of \(J^*\) in comparison to \(V^*\) when going towards the thermodynamic limit.}
	\label{J_bandwidth}
\end{figure}

Since the high momentum modes have been decoupled, the ground state in this regime can be estimated by compressing the kinetic energy to the zero bandwidth limit: 
\begin{equation}\begin{aligned}
	\sum_{\sigma,\vec k:|\epsilon_{\vec k}| < D^*} \epsilon_{\vec k} \tau_{\vec k,\sigma} \sim \epsilon_F \sum_{\vec k_F,\sigma}\tau_{\vec k_F,\sigma}
\end{aligned}\end{equation}
Choosing the chemical potential such that \(\epsilon_F\) vanishes results in a {\it two-site Hamiltonian} composed of the impurity site and the zeroth site (see Fig.~\ref{zero-bw}):
\begin{equation}
	\label{two-site}
	H_\text{eff}^\text{sc} \sim J^* \vec{S}_d\cdot\vec{S}_0 - U_b \hat n_{0 \uparrow} \hat n_{0 \downarrow}
\end{equation}
\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.4\textwidth]{../figures/zeromode_eff.pdf}
	\caption{Two site Hamiltonian obtained by taking the zero bandwidth limit of the effective Hamiltonian in the screened regime.}
	\label{zero-bw}
\end{figure}
This model is easily solved and the ground state is a maximally-entangled spin singlet formed by \(\vec S_d\) and \(\vec S_0\); the ground state energy is \(E_\text{gs}^\text{sc} = -\frac{3}{4}J^* - U_b\):
\begin{equation}\begin{aligned}
	\ket{\Psi}_\text{gs}^\text{sc} = \frac{1}{\sqrt 2}\left(\ket{\uparrow_d,\downarrow_0} - \ket{\downarrow_d,\uparrow_0}\right)
\end{aligned}\end{equation}
where \(\left\{\uparrow_d, \downarrow_d\right\}, \left\{\uparrow_0, \downarrow_0\right\}\) are the spin states of the impurity and zeroth sites respectively. As mentioned before, this regime in the GAIM is adiabatically connected to the AIM, and indeed, such a spin-singlet ground state is also obtained in the AIM from other methods like Poor Man's scaling~\cite{anderson1969exact,anderson1970}, numerical renormalisation group~\cite{wilson1975,hrk_wilson_1980}, Bethe ansatz~\cite{}, conformal field theory methods~\cite{}[CITE] and the URG analysis of the Kondo model~\cite{anirban_kondo}.

@@

\begin{figure}[htpb]
	\centering
\includegraphics[width=0.5\textwidth]{../figures/corrs_gs.pdf}
\caption{Overlap of the RG fixed point ground state \(\ket{\Psi}_\text{gs}\) with the spin singlet state \(\ket{\text{ss}}\) and the charge triplet state \(\ket{\text{ct}}\), for a range of values of the transition tuning ratio \(r\). Beyond the critical point \(r > 0.25\), both these overlaps vanish.}
	\label{overlaps}
\end{figure}

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.5\textwidth]{../figures/Add.pdf}
	\caption{Impurity spectral function \(A_d(\omega)\) at three values of the ratio \(r\). The central large plot is exactly at the critical point \(r=0.25\), the left (blue) inset is at \(r < 0.25\) and the right (violet) inset is at \(r > 0.25\).}
	\label{spec-func}
\end{figure}

@@

\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.5\textwidth]{../figures/entanglement.pdf}
	\caption{Variation of \(\sqrt{\left(1 - \varepsilon_\text{ss} \right)}\sqrt{\left(1 - \varepsilon_\text{ct} \right)}\) as \(r\) is increased towards the critical value 0.25. Inset shows the variation of the individual measures \(\varepsilon_\text{ss},\varepsilon_\text{ct}\).}
	\label{entng}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=0.49\textwidth]{../figures/I_k.pdf}
\caption{Mutual information in \(k-\)space, across the transition. Red curve is the mutual information between two distinct \(k-\) states inside the Kondo cloud. Blue curve is the mutual information between a \(k-\)state and the impurity state of opposite spin.}
\label{mut_inf_k}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=0.49\textwidth]{../figures/I_r.pdf}
\caption{Mutual information in \(r-\)space, across the transition. Red curve is the mutual information between the zeroth site and the site next to it. Blue curve is the mutual information between a \(k-\)state and the impurity state of opposite spin. The former is zero before the transition because the zeroth site is locked in a singlet with the impurity, and becomes zero after  the transition because the singlet has now decoupled and the zeroth site is free to entangle with the other sites.}
\label{mut_inf_r}
\end{figure}

>>>>

The Hubbard model has become one of the fundamental and prototypical models for describing strongly correlated systems in quantum condensed matter physics.   strong electronic correlation; in its simplest form, it features a single band of conduction electrons hopping on a lattice and interacting via local correlations that provide a cost \(U\) if any site is doubly occupied. There are two trivial limits of the model. At \(U=0\), the bath consists of just a kinetic energy part, and the ground state is just a filled Fermi sea. At \(t=0\), each lattice site decouples from the rest and becomes a local moment, which under symmetry-breaking becomes a Neel antiferromagnet. This suggests that on increasing \(U/t\) beyond some critical value, the system might undergo a phase transition from a metallic state to an insulating state~\cite{Mott_1949}. This transition is reflected in the local spectral function - while it has a well-defined zero energy peak in the metallic phase, it is gapped in the insulating phase.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.5\textwidth]{dos_diff.pdf}
\caption{Two kinds of bath that an impurity can hybridise into. The left panel shows a non-interacting conduction band with a flat density of states. The right panel shows an interacting bath with an energy-dependent density of states. In the latter case, the impurity "feels" a reduced effective bandwidth defined by the width of the central peak.}
\end{figure}

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.45\textwidth]{zeromode_bare.pdf}
	\caption{While we have studied the full model under renormalisation group, often we will turn to a simplified zero-bandwidth version of the model that is obtained by ignoring the kinetic energy part of the Hamiltonian. This zero-bandwidth model is effectively a two site model.}
	\label{zeromode-bare}
\end{figure}

@@

\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.45\textwidth]{urg-scheme.pdf}
	\caption{High energy - low energy scheme defined and used in the URG method. The states away from the Fermi surface form the UV subspace and are decoupled first, leading to a Hamiltonian which is more block-diagonal and comprised of only the IR states near the Fermi surface.}
	\label{urg-scheme}
\end{figure}

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.48\textwidth]{J_Ub.pdf}
	\includegraphics[width=0.48\textwidth]{U_Ub.pdf}
	\includegraphics[width=0.48\textwidth]{V_Ub.pdf}
	\caption{Variation of couplings \(U,V\) and \(J\) along the RG transformations, for three values of the transition-tuning ratio \(r\). x-axis represents the RG steps, the rightmost point is the first RG step and leftmost point is the final RG step. The values of the bare couplings are given in the title. All values are in units of the half-bandwidth \(D_0\), which is chosen to be 10. The blue curves represent the flows for \(r > r_c\), where \(J\) is relevant. The purple curves represent the RG flows for \(r < r_c\), where \(J\) is irrelevant. The red curves are exactly at the critical point \(r = r_c\), where \(J\) is marginal.}
	\label{rg-flow}
\end{figure}

@@

\begin{figure}[!htb]
	\includegraphics[width=0.5\textwidth]{phase-map-MIT.pdf}
	\caption{Phase diagram of the GAIM, in the space of couplings \(U_0\) and \(J_0\). \(U_b\) is set to \(-U_0/10\) and \(V_0\) is set to \(2J_0\). The various phases \(A\) through \(D\) are characterised in the text.}
	\label{phase-diag}
\end{figure}

@@

\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.49\textwidth]{odlro.pdf}
	\caption{Growth in the pair-fluctuation terms \(c^\dagger_{k \uparrow}c^\dagger_{k \downarrow} c_{k^\prime \downarrow}c_{k^\prime \uparrow}\), as \(r \to 0.25\). This comes at the cost of the spin-flip fluctuations \(S_d^+ S_0^- + \text{h.c.}\), which decrease towards the transition. The inset plots the same quantities but for a larger range of \(r\), and the pair fluctuations are seen to pick up abruplty after the transition while the spin fluctuations vanish.}
	\label{pair_fluc}
\end{figure}

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.5\textwidth]{J_bandwidth.pdf}
	\caption{Increase of \(J^*\) in comparison to \(V^*\) when going towards the thermodynamic limit.}
	\label{J_bandwidth}
\end{figure}

Since the high momentum modes have been decoupled, the ground state in this regime can be estimated by compressing the kinetic energy to the zero bandwidth limit: 
\begin{equation}\begin{aligned}
	\sum_{\sigma,\vec k:|\epsilon_{\vec k}| < D^*} \epsilon_{\vec k} \tau_{\vec k,\sigma} \sim \epsilon_F \sum_{\vec k_F,\sigma}\tau_{\vec k_F,\sigma}
\end{aligned}\end{equation}
Choosing the chemical potential such that \(\epsilon_F\) vanishes results in a {\it two-site Hamiltonian} composed of the impurity site and the zeroth site (see Fig.~\ref{zero-bw}):
\begin{equation}
	\label{two-site}
	H_\text{eff}^\text{sc} \sim J^* \vec{S}_d\cdot\vec{S}_0 - U_b \hat n_{0 \uparrow} \hat n_{0 \downarrow}
\end{equation}
\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.4\textwidth]{zeromode_eff.pdf}
	\caption{Two site Hamiltonian obtained by taking the zero bandwidth limit of the effective Hamiltonian in the screened regime.}
	\label{zero-bw}
\end{figure}
This model is easily solved and the ground state is a maximally-entangled spin singlet formed by \(\vec S_d\) and \(\vec S_0\); the ground state energy is \(E_\text{gs}^\text{sc} = -\frac{3}{4}J^* - U_b\):
\begin{equation}\begin{aligned}
	\ket{\Psi}_\text{gs}^\text{sc} = \frac{1}{\sqrt 2}\left(\ket{\uparrow_d,\downarrow_0} - \ket{\downarrow_d,\uparrow_0}\right)
\end{aligned}\end{equation}
where \(\left\{\uparrow_d, \downarrow_d\right\}, \left\{\uparrow_0, \downarrow_0\right\}\) are the spin states of the impurity and zeroth sites respectively. As mentioned before, this regime in the GAIM is adiabatically connected to the AIM, and indeed, such a spin-singlet ground state is also obtained in the AIM from other methods like Poor Man's scaling~\cite{anderson1969exact,anderson1970}, numerical renormalisation group~\cite{wilson1975,hrk_wilson_1980}, Bethe ansatz~\cite{}, conformal field theory methods~\cite{}[CITE] and the URG analysis of the Kondo model~\cite{anirban_kondo}.

@@

\begin{figure}[htpb]
	\centering
\includegraphics[width=0.5\textwidth]{corrs_gs.pdf}
\caption{Overlap of the RG fixed point ground state \(\ket{\Psi}_\text{gs}\) with the spin singlet state \(\ket{\text{ss}}\) and the charge triplet state \(\ket{\text{ct}}\), for a range of values of the transition tuning ratio \(r\). Beyond the critical point \(r > 0.25\), both these overlaps vanish.}
	\label{overlaps}
\end{figure}

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.5\textwidth]{Add.pdf}
	\caption{Impurity spectral function \(A_d(\omega)\) at three values of the ratio \(r\). The central large plot is exactly at the critical point \(r=0.25\), the left (blue) inset is at \(r < 0.25\) and the right (violet) inset is at \(r > 0.25\).}
	\label{spec-func}
\end{figure}

@@

\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.5\textwidth]{entanglement.pdf}
	\caption{Variation of \(\sqrt{\left(1 - \varepsilon_\text{ss} \right)}\sqrt{\left(1 - \varepsilon_\text{ct} \right)}\) as \(r\) is increased towards the critical value 0.25. Inset shows the variation of the individual measures \(\varepsilon_\text{ss},\varepsilon_\text{ct}\).}
	\label{entng}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=0.49\textwidth]{I_k.pdf}
\caption{Mutual information in \(k-\)space, across the transition. Red curve is the mutual information between two distinct \(k-\) states inside the Kondo cloud. Blue curve is the mutual information between a \(k-\)state and the impurity state of opposite spin.}
\label{mut_inf_k}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=0.49\textwidth]{I_r.pdf}
\caption{Mutual information in \(r-\)space, across the transition. Red curve is the mutual information between the zeroth site and the site next to it. Blue curve is the mutual information between a \(k-\)state and the impurity state of opposite spin. The former is zero before the transition because the zeroth site is locked in a singlet with the impurity, and becomes zero after  the transition because the singlet has now decoupled and the zeroth site is free to entangle with the other sites.}
\label{mut_inf_r}
\end{figure}
